---
title: "AE 01: Bechdel + data visualization"
author: "Insert name"
format: html
editor: visual
filters: 
  - webr
webr:
  packages: ['tidyverse', 'scales']
---

```{webr-r}
#| context: setup
download.file(
  "https://raw.githubusercontent.com/ElijahMeyer3/Coursera/main/data/bechdel.csv",
  "bechdel.csv"
)
```

```{r}
#| echo: false
#| context: setup

download.file(
  "https://raw.githubusercontent.com/ElijahMeyer3/Coursera/main/data/bechdel.csv",
  "bechdel.csv"
)
```

```{r}
#| echo: false
bechdel <- read.csv("bechdel.csv")
```


In this mini analysis we work with the data used in the FiveThirtyEight story titled ["The Dollar-And-Cents Case Against Hollywood's Exclusion of Women"](https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/).

This analysis is about the [Bechdel test](https://en.wikipedia.org/wiki/Bechdel_test), a measure of the representation of women in fiction.

## Getting started

### Packages

We start with loading the packages we'll use: **tidyverse** for majority of the analysis and **scales** for pretty plot labels later on.

```{webr-r}
#| label: load-packages
#| warning: false

library(tidyverse)
library(scales)
```

### Data

The data are stored as a CSV (comma separated values) file in the `data` folder of your repository.
Let's read it from there and save it as an object called `bechdel`.



```{webr-r}
#| label: load-data
bechdel <- read.csv("bechdel.csv")
```

::: callout-note
This a modified version of the `bechdel` dataset from the previous application exercise.
It's been modified to include some new variables derived from existing variables as well as to limit the scope of the data to movies released between 1990 and 2013.
For now we're not going to discuss how these modifications were made (that's next week's topic) but if you're curious, you can find the data prep script in the `data/` folder of your repo.
Don't spend too long trying to decipher it, wait till next week when you have the right tools to do so!
:::

### Get to know the data

We can use the `glimpse` function to get an overview (or "glimpse") of the data.

```{webr-r}

# add code here

```

With your output, confirm that: 

-- There are  movies in the dataset

-- There are  variables (columns) in the dataset


What does each observation (row) in the data set represent?

<label>Answer:</label>
<br>
<textarea rows="4" cols="50">
</textarea><br><br>


::: {.callout-tip collapse="true"}
## Solution
Each observation represents a **movie**.
:::

## Variables of Interest 

The variables we'll focus on are the following:

-   `budget_2013`: Budget in 2013 inflation adjusted dollars.
-   `gross_2013`: Gross (US and international combined) in 2013 inflation adjusted dollars.
-   `roi`: Return on investment, calculated as the ratio of the gross to budget.
-   `clean_test`: Bechdel test result:
    -   `ok` = passes test
    -   `dubious`
    -   `men` = women only talk about men
    -   `notalk` = women don't talk to each other
    -   `nowomen` = fewer than two women
-   `binary`: Bechdel Test PASS vs FAIL binary

We will also use the `year` of release in data prep and `title` of movie to take a deeper look at some outliers.

There are a few other variables in the dataset, but we won't be using them in this analysis.

## Visualizing data with `ggplot2`

**ggplot2** is the package and `ggplot()` is the function in this package that is used to create a plot.

-   `ggplot()` creates the initial base coordinate system, and we will add layers to that base. We first specify the data set we will use with `data = bechdel`.

```{webr-r}
#| label: plot-base
ggplot(data = bechdel)
```

-   The `mapping` argument is paired with an aesthetic (`aes()`), which tells us how the variables in our data set should be mapped to the visual properties of the graph.

```{webr-r}
#| label: plot-aesthetic-mapping
ggplot(data = bechdel, 
       mapping = aes(x = budget_2013, y = gross_2013))
```

As we previously mentioned, we often omit the names of the first two arguments in R functions.
So you'll often see this written as:

```{webr-r}
#| label: plot-simplified-call
ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013))
```

Note that the result is exactly the same.

-   The `geom_xx` function specifies the type of plot we want to use to represent the data. In the code below, we use `geom_point` which creates a plot where each observation is represented by a point.

```{webr-r}
#| label: plot-point
ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013)) +
       # type geom_point() here!
```



::: {.callout-tip collapse="true"}
## Expected Result

```{r}
#| echo: false
library(ggplot2)
ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013)) +
       geom_point()
```
:::

Note that this results in a warning as well. 

This warning represents the number of observations that were removed because there were missing data!

## Gross revenue vs. budget

### Step 1 - Your turn

The following code changes the color of all points to coral. Explore different colors by changing "coral" to different colors!

::: callout-tip
See <http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf> for many color options you can use by name in R or use the [hex code](https://en.wikipedia.org/wiki/Web_colors) for a color of your choice.
:::

```{webr-r}
#| label: plot-point-color

ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013)) +
  geom_point(color = "coral") 
```

### Step 2 - Your turn

Add labels for the title and x and y axes using [labs](https://ggplot2.tidyverse.org/reference/labs.html). Do this by modifying the existing code below.

```{webr-r}
#| label: plot-labs

ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013)) +
  geom_point(color = "coral") + 
  # insert code here
```

::: {.callout-tip collapse="true"}
## Expected Result
```{r}
#| label: plot-labels

ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013))+
  geom_point(color = "deepskyblue3") + 
  labs(
    x = "Budget (in 2013 $)", 
    y = "Gross revenue (in 2013 $)", 
    title = "Gross revenue vs. budget"
    )
```
:::

### Step 3 - Your turn

An `aes`thetic is a visual property of one of the objects in your plot.
Commonly used aesthetic options are:

-   color
-   fill
-   shape
-   size
-   alpha (transparency)

Modify the plot below, so the `color` of the points is based on the variable `binary`, and make the size of your points based on `roi`. 


```{webr-r}
#| label: plot-aes-color

ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013)) +
  geom_point() + 
  labs(
    x = "Budget (in 2013 $)", 
    y = "Gross revenue (in 2013 $)", 
    title = "Gross revenue vs. budget, by Bechdel test result"
    )
```

::: {.callout-tip collapse="true"}
## Expected Result
```{r}
#| label: plot-aes-color-solution

ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013,
           color = binary, size = roi)) +
  geom_point() + 
  labs(
    x = "Budget (in 2013 $)", 
    y = "Gross revenue (in 2013 $)", 
    title = "Gross revenue vs. budget, by Bechdel test result"
    )
```
:::

### Step 4 - Your turn

Expand on your plot from the previous step to make the transparency (`alpha`) of the points 0.5.

```{webr-r}
#| label: plot-aes-alpha

ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013,
           color = binary, size = roi)) +
  geom_point(alpha = #insert 0.5 here) + 
  labs(
    x = "Budget (in 2013 $)", 
    y = "Gross revenue (in 2013 $)", 
    title = "Gross revenue vs. budget, by Bechdel test result"
    )
```

### Step 5 - Your turn

Expand on your plot from the previous step by using `facet_wrap` to display the association between budget and gross for different values of `clean_test`.

```{webr-r}
#| label: plot-facet-wrap

ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013,
           color = binary, size = roi)) +
  geom_point(alpha = 0.5) + 
  facet_wrap(~#insert variable here) +
  labs(
    x = "Budget (in 2013 $)", 
    y = "Gross revenue (in 2013 $)", 
    title = "Gross revenue vs. budget, by Bechdel test result"
    )
```

### Step 6 - Demo

Improve your plot from the previous step by making the x and y scales more legible.

::: callout-tip
Make use of the [**scales**](https://scales.r-lib.org/) package, specifically the `scale_x_continuous()` and `scale_y_continuous()` functions.
:::

```{r}
#| label: plot-scales

library(scales)

ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013,
           color = binary, size = roi)) +
  geom_point(alpha = 0.5) + 
  facet_wrap(~clean_test) +
  scale_x_continuous(labels = label_dollar(scale = 1/1000000)) +
  scale_y_continuous(labels = label_dollar(scale = 1/1000000)) +
  labs(
    x = "Budget (in 2013 $)", 
    y = "Gross revenue (in 2013 $)", 
    title = "Gross revenue vs. budget, by Bechdel test result"
    )
```

### Step 7 - Your turn

Expand on your plot from the previous step by using `facet_grid` to display the association between budget and gross for different combinations of `clean_test` and `binary`.
Comment on whether this was a useful update.

```{webr-r}
#| label: plot-facet-grid

ggplot(bechdel, 
       aes(x = budget_2013, y = gross_2013,
           color = binary, size = roi)) +
  geom_point(alpha = 0.5) + 
  facet_grid(#insert variable name1 ~ insert variable name2) +
  scale_x_continuous(labels = label_dollar(scale = 1/1000000)) +
  scale_y_continuous(labels = label_dollar(scale = 1/1000000)) +
  labs(
    x = "Budget (in 2013 $)", 
    y = "Gross revenue (in 2013 $)", 
    title = "Gross revenue vs. budget, by Bechdel test result"
    )
```

Is this type of facet useful? Why or why not?

<label>Answer:</label>
<br>
<textarea rows="4" cols="50">
</textarea><br><br>

::: {.callout-tip collapse="true"}
## Solution
This was not a useful update as one of the levels of `clean_test` maps directly to one of the levels of `binary`.
:::

## Step 8 - Demo 

What other improvements could we make to this plot?

<label>Answer:</label>
<br>
<textarea rows="4" cols="50">
</textarea><br><br>

::: {.callout-tip collapse="true"}
## Solution
We could make the plot better by... 

- 

- 

-
:::

```{webr-r}
#| label: plot-improve

# Answers may vary
```


## Assessment Questions 


 